A user-extensible and adaptable parser architecture

### 2008-06-01 papers/sgai-2008/paper.tex
Shaved another half a page off: * Removed the glossary entirely, moving the SMTP glossary entry to the introduction * Reduced the introduction by several lines, to fit in the SMTP footnote and   pull the start of the framework/actions/rules list onto page 2. This should give me a bit of space for an example action, rule and log line (or two).

\usepackage[style=altlist,number=none,toc=true,global=true]{glossary}
\makeglossary{}

@@

\newacronym{ATN}{Augmented Transition Networks}{description={
    Augmented Transition Networks, originally described in~\cite{atns} and
    further in~\cite{nlpip}, are a tool used in Computational Linguistics
    for creating grammars to parse or generate sentences,
}}

\newacronym{DNSBL}{DNS Blacklist}{description={
    A DNS Blacklist is a simple collaborative anti-spam technique used to
    reject or penalise email sent from mail servers reported to have sent
    large volumes of spam.
}}

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP is the protocol which transfers mail from the sender to the
    recipient across the Internet.  It is a simple, human readable, plain
    text protocol, making it quite easy to test and debug problems with it.
    A detailed description of \SMTP{} is beyond the scope of this document:
    the original protocol definition is RFC~821~\cite{RFC821}, updated in
    RFC~2821~\cite{RFC2821}.
}}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing new measures to supplement the
existing filters where there is a deficiency.  Most anti-spam techniques
are content-based~\cite{a-plan-for-spam, word-stemming,
relaxed-online-svms}, viz they examine the contents of each mail to
determine whether it is spam or not.  Examining content requires accepting
the mail in the first place, but rejecting rather than accepting mail is
preferable in many cases: senders of non-spam mail that is mistakenly
rejected will receive an immediate non-delivery notice; load is reduced on
mail servers (allowing more intensive content-based measures to be used on
mail that is accepted); users have less mail to wade through.  Improving
the performance of anti-spam measures which are applied when mail is being
transferred via \SMTP{} is the goal of this project, by providing a
platform for reasoning about anti-spam filters.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The parser must be
designed so that adding support for parsing new log lines is a simple task,
as the log lines will continue to evolve over time.  The variety in log
lines occurs for several reasons:

@@

    \item The mail administrator can define custom rejection
        messages, resulting in custom log lines.

    \item External resources utilised by Postfix (e.g.\ policy
        servers~\cite{policy-servers} or \DNSBL{} lookups) can also change
        their messages, sometimes without warning.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.  The effort required to adapt and improve an
existing parser was judged to be greater than the effort to write one from
scratch, as the techniques used by the existing parsers severely limited
their potential: some ignored the majority of log lines, parsing specific
log lines with great accuracy, but in an inextensible fashion; others
sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same type, e.g.\ rejections, and
extracted only a portion of the data required for this project.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific
\DNSBL{}).  This architecture is ideally suited to parsing inputs where the
input is not fully understood or does not follow a fixed grammar: the
architecture warns about unparsed inputs and other errors, but continues on
parsing as best it can, allowing the developer of a new parser to decide
which deficiencies are most important and require attention first, rather
than being forced to fix the first error which arises.

@@

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is interesting how two
different approaches have a similar division of responsibilities, while
having completely different implementations and semantics.

@@

\begin{table}[ht]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \hline
        \noalign{\smallskip}
        \ATN{}        & Parser    & Similarity                          \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
        Networks      & Framework & Determines the sequence of
                                    transitions or actions which        \\
                      &           & constitutes a valid input.          \\
        Transitions   & Actions   & Save data and impose conditions the
                                    input must meet to be               \\
                      &           & considered valid.                   \\
        Abbreviations & Rules     & Responsible for classifying input.  \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
    \end{tabular}
\end{table}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \SMTP{} delivery to specific
sites from \SMTP{} delivery to the rest of the world.  The algorithm
provides a facility for specifying the order of overlapping rules: the
priority field in each rule (defaults to zero).  Rules are sorted by
priority, highest first, and then rules with the same priority are sorted
by the number of successful matches when parsing the previous log file.
Negative priorities may be useful for catchall rules.

@@

\bibliographystyle{spmpsci}
\bibliography{../common/bibliography}
\label{bibliography}

% Redefine the command used to produce the glossary title so that it
% doesn't produce a new section - this will save space.
\renewcommand{\glossarytitle}{\noindent{}\textbf{\large{Glossary}}}
\printglossary{}

>>>>

@@

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental measures where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting is
preferable in many cases: senders of non-spam mail that is mistakenly
rejected will receive an immediate non-delivery notice; load is reduced on
mail servers (allowing more intensive content-based measures to be used);
users have less mail to wade through.  Improving the performance of
anti-spam measures which are applied when mail is being transferred via
SMTP\footref{smtp footnote} is the goal of this project, by providing a
platform for reasoning about anti-spam filters.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the SMTP\footnote{\label{smtp footnote}Simple Mail
Transfer Protocol transfers mail across the Internet.  It is a simple,
human readable, plain text protocol, making it quite easy to test and debug
problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.} server in use.
The parser must be designed so that adding support for parsing new log
lines is a simple task, as the log lines will continue to evolve over time.
The variety in log lines occurs for several reasons:

@@

    \item The mail administrator can define custom rejection
        messages.

    \item External resources utilised by Postfix (e.g.\ policy
        servers~\cite{policy-servers} or DNSBLs) can also change their
        messages, sometimes without warning.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data.
The effort required to adapt and improve an existing parser was judged to
be greater than the effort to write one from scratch, as the techniques
used by the existing parsers severely limited their potential: some ignored
the majority of log lines, parsing specific log lines with great accuracy,
but in an inextensible fashion; others sloppily parsed the majority of log
lines, but were incapable of distinguishing between log lines of the same
type, e.g.\ rejections, and extracted only a portion of the data required.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejections), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type, and only that type (e.g.\ rejection by a specific DNSBL).
This architecture is ideally suited to parsing inputs where the input is
not fully understood or does not follow a fixed grammar: the architecture
warns about unparsed inputs and other errors, but continues on parsing as
best it can, allowing the developer of a new parser to decide which
deficiencies are most important and require attention first, rather than
being forced to fix the first error which arises.

@@

There is some similarity between the parser's design and William Wood's
Augmented Transition Networks (ATN)~\cite{atns, nlpip}, a tool used in
Computational Linguistics for creating grammars to parse or generate
sentences.  The resemblance between ATN and the parser is accidental, but
it is interesting how two different approaches have a similar division of
responsibilities, while having completely different implementations and
semantics.

@@

\begin{table}[ht]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \hline
        \noalign{\smallskip}
        ATN           & Parser    & Similarity                          \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
        Networks      & Framework & Determines the sequence of
                                    transitions or actions which        \\
                      &           & constitutes a valid input.          \\
        Transitions   & Actions   & Save data and impose conditions the
                                    input must meet to be               \\
                      &           & considered valid.                   \\
        Abbreviations & Rules     & Responsible for classifying input.  \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
    \end{tabular}
\end{table}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  The algorithm provides
a facility for specifying the order of overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches when parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

\bibliographystyle{spmpsci}
\bibliography{../common/bibliography}
